HMM Model

### 2020-10-23 manuscript/analysis_brms.Rmd
Cleanups, tests, more complex HMM models

---
title: "Analysis of Retrospective COVID-19 data"
author: "Paul Bürkner"
output:
  html_document:
    theme: default
encoding: UTF-8
editor_options: 
  chunk_output_type: console
---

@@

```{r}
draws <- tidy_draws(fit_discharged_time)
hypothesis_res_list[["hcq_az_discharged_time"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiraviTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$favipiravi_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

>>>>

---
title: "Analysis of Retrospective COVID-19 data"
author: "Paul Bürkner"
output:
  html_document:
    theme: default
encoding: UTF-8
#editor_options: 
#  chunk_output_type: console
---

@@

```{r}
draws <- tidy_draws(fit_discharged_time)
hypothesis_res_list[["hcq_az_discharged_time"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### 2020-10-23 manuscript/hmm.Rmd
Cleanups, tests, more complex HMM models

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care),
         is_male = as.integer(sex == "M")
         )
  #mutate(best_supportive_care = as.integer(best_supportive_care))

serie_data_28 <- wide %>%
  inner_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

# Extend terminal states for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  ungroup()

@@

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_group_m1_counterfactual_took_hcq <- trt_group_m1
# trt_group_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_group_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_took_hcq)
# 
# trt_group_m1_counterfactual_no_hcq <- trt_group_m1
# trt_group_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_group_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_no_hcq)

@@

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

@@

```{r trt_rate_m2}
trt_rate_m2_formula <- update.formula(trt_rate_m1_formula,  ~ . +  (0 + age_norm + is_male || rate_group))
trt_rate_m2 <- brmshmmdata(trt_rate_m2_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

@@

```
```{r}
do_common_pp_checks(c_trt_m1_fit)
```

>>>>

```{r}
transform_serie_data <- function(raw_serie_data) {
  raw_serie_data  %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care),
         is_male = as.integer(sex == "M")
         )
}

serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

serie_data_28 <- wide %>%
  right_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

# Extend terminal states and took_XX data for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  mutate(across(tidyselect::starts_with("took_"), ~ if_else(is.na(.x), as.integer(any(.x != 0, na.rm = TRUE)), .x))) %>%
  ungroup()

@@

```{r}
# fit <- trt_group_m1_fit
# data_28 <- trt_group_m1
# data_28$serie_data <- serie_data_28
# epred_rect <- posterior_epred_rect(fit)#, newdata = data_28)
# predicted_rect <- posterior_epred_to_predicted(fit, epred_rect)
# # predicted <- posterior_rect_to_long(fit, predicted_rect)
# 
# pp_check_state_at(fit, predicted_rect, 28, newdata = data_28)
```

```{r}
# counterfactual_took_hcq <- serie_data_28 %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_group_m1_counterfactual_took_hcq <- trt_group_m1
# trt_group_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_group_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_took_hcq)
# 
# trt_group_m1_counterfactual_no_hcq <- trt_group_m1
# trt_group_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_group_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_no_hcq)

@@

```{r}
# fit <- trt_group_m1_fit
# epred_rect <- posterior_epred_rect(fit)
# predicted_rect <- posterior_epred_to_predicted(fit, epred_rect)
# predicted <- posterior_rect_to_long(fit, predicted_rect)

if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

@@

```{r trt_rate_m2}
trt_rate_m2_formula <- update.formula(trt_rate_m1_formula,  ~ . +  (0 + age_norm + is_male || .rate_id))
trt_rate_m2 <- brmshmmdata(trt_rate_m2_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

@@

```
```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m1_fit)
}
```

```{r c_trt_m2}
c_trt_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m2_fit <- brmhmm(c_trt_m2, cache_file = paste0(cache_dir,"/c_trt_m2.rds"), init = 0.1, control = default_control, iter = 1200)

#parnames(c_trt_m2_fit$brmsfit)
summary(c_trt_m2_fit$brmsfit)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m2_fit)
}
```

```{r c_trt_m3}
c_trt_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m3_fit <- brmhmm(c_trt_m3, cache_file = paste0(cache_dir,"/c_trt_m3.rds"), init = 0.1, control = default_control, iter = 1200)

summary(c_trt_m3_fit$brmsfit)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m3_fit)
}
```

```{r c_trt_m4}
c_trt_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m4_fit <- brmhmm(c_trt_m4, cache_file = paste0(cache_dir,"/c_trt_m4.rds"), init = 0.1, control = default_control, iter = 1200)

summary(c_trt_m4_fit$brmsfit)
```

```{r}
# draws <- tidybayes::tidy_draws(c_trt_m4_fit$brmsfit)
# hmm_hypothesis_res_list[["c_trt_m4_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
# 
# hmm_hypothesis_res_list[["c_trt_m4_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m4_fit)
}
```
